Domain-specific working memory loads selectively increase negative interpertations of surprised facial expressions

### 2019-09-06 Manuscript/Manuscript.Rmd
Cleaned up scripts
Made a MERP specific data cleaning script...

Starting to polish methods & results.

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

@@

# Methods
## Participants
58 subjects were recruited from the University of Nebraska-Lincoln. All subjects provided written informed consent in accordance with the Declaration of Helsinki and all procedures were approved by the local Institutional Review Board (Approval #20141014670EP). The data from eight subjects were excluded due to technical difficulties or an error in the experiment script. This left 50 individuals in the final sample.

## Material
### Stimuli
``` {r stimulus_t.test, include = FALSE}
stim.data <- read.xlsx("~/Documents/Nick-Grad/Neta_Lab/depletion_study/study2/Analyses/WorkingMemoryLoads/IAPS_Stim_List.xlsx")
shapiro.test(subset(stim.data, Condition == "POS")$Aro_Mn)
shapiro.test(subset(stim.data, Condition == "NEG")$Aro_Mn)
stim.t.test <- wilcox_test(subset(stim.data, Condition == "POS")$Aro_Mn, subset(stim.data, Condition == "NEG")$Aro_Mn)
```
The stimuli included faces taken from the NimStim [@tottenham_nimstim_2009] and Karolinska Directed Emotional Faces [@lundqvist_karolinska_1998] stimuli sets. The faces consisted of 34 unique identities including 11 angry, 12 happy, and 24 surprised expressions organized pseudorandomly. The scenes were taken from the International Affective Picture System [@lang_international_2008]. A total of 288 scenes (72 positive, 72 negative, and 144 neutral) were selected for the image matrices. The positive and negative images did not differ on arousal (Z = `r printnum(stim.t.test$z_val)`, p = `r printnum(stim.t.test$p.value)`). The scenes were organized into low (two images) and high (six images) cognitive load of either neutral or emotional (equal number of positive and negative) images (Figure 1).

@@

## Procedure
Participants were randomly assigned to complete one of the task versions. The tasks included 144 [^1] trials split between working memory probe and face rating trials. On each trial, participants first viewed an image matrix of either neutral or emotional images, which the participants were instructed to remember for the duration of the trial.  The image matrix was presented for four seconds and the images were swapped from low and high load matrices across versions of the task. After the image matrix a happy, angry, or surprised face appeared for one second and the participants rated the face by clicking on either the positive or negative response option. After the face rating, a single image probe appeared, and participants indicated whether or not the image probe was present in the previous image matrix.

@@

# Results
## Subjective ratings
```{r import data and analyze ratings, include = FALSE, echo = TRUE}
data <- read.csv("~/Documents/Nick-Grad/Neta_Lab/depletion_study/study2/Analyses/WorkingMemoryLoads/Data/Cleaned_Data/Final.Data.csv_2019-08-14_08-52-46.csv")
### remove psych disorders ### Ask maital?
# data <- data[-c(4, 33, 37, 38, 41, 47),]

@@

## Reaction times
```{r analyze RTs, include = FALSE}
## assess normality ###
shapiro.test(data$lo.neu.sur_RT)
shapiro.test(data$hi.neu.sur_RT)
shapiro.test(data$lo.emo.sur_RT)
shapiro.test(data$hi.emo.sur_RT)

### make long ###
data.long <- gather(data, key = "Condition", value = "ReactionTime",
                    lo.neu.sur_RT, hi.neu.sur_RT,
                    lo.emo.sur_RT, hi.emo.sur_RT)

### make factors ###
data.long$Load <- ifelse(data.long$Condition == "lo.neu.sur_RT", "Low",
                         ifelse(data.long$Condition == "lo.emo.sur_RT", "Low",
                                ifelse(data.long$Condition == "hi.neu.sur_RT", "High",
                                       ifelse(data.long$Condition == "hi.emo.sur_RT", "High", ""))))

data.long$Type <- ifelse(data.long$Condition == "lo.neu.sur_RT", "Neutral",
                         ifelse(data.long$Condition == "lo.emo.sur_RT", "Emotional",
                                ifelse(data.long$Condition == "hi.neu.sur_RT", "Neutral",
                                       ifelse(data.long$Condition == "hi.emo.sur_RT", "Emotional", ""))))

### make w/in subjs factors ###
data.long$subjID <- as.factor(data.long$subjID)

### traditional ANOVA ###
RTaov <- with(data.long,
                   aov(ReactionTime ~ (Load * Type) +
                       Error(subjID / (Load * Type))), contrasts = contr.sum())
summary(RTaov)
RTaov <- tidy(RTaov)
```
To assses for normality, the data were first tested with Shapiro-Wilk's test. All conditions appeared to be sampled from normal distributions (p's > .300). A Load (Low, High) X Type (Neutral, Emotional) repeated measures ANOVA was used to assess differences in the RTs. There was a trend towards a main effect of domain, F(`r (RTaov$df[which(RTaov$term == "Type")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Type")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Type")])`. There was no effect of load, F(`r (RTaov$df[which(RTaov$term == "Load")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Load")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Load")])`, nor any interaction between load and domain, F(`r (RTaov$df[which(RTaov$term == "Load:Type")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Load:Type")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Load:Type")])`.

```{r plot RT data, include = TRUE}
ggplot(data.long, aes(Condition, `ReactionTime`, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)
```
Next, we tested for reaction time differences between positive and negative interpretations of surprised facial expressions within each working memory load condition. A large proportion of the conditions were not normally distributed, thus we used paired-sample Wilcoxon signed rank tests for assessing differences between positive and negative ratings of surprise RTs. This difference was significant *only* for the low emotional load trials, with surprise rated as positive taking signficantly longer than surprise rated as negative (`r printp(lo.neu.RT.pvn$p.value)`). All other comparisons were not significant (p's > .200). Further, this effect survived Bonferroni correction (p = `r printp(.05 / 4)`).

>>>>

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

# Methods
## Participants
``` {r read data}
data <- read.csv("~/Documents/Nick-Grad/Neta_Lab/depletion_study/study2/Analyses/WorkingMemoryLoads/Data/Cleaned_Data/Final.Data.csv_2019-09-06_12-18-56.csv")
```
Fifty-eight subjects were recruited from the undergraduate research pool at the University of Nebraska-Lincoln. The data from eight subjects were excluded due to technical difficulties resulting from an error in one of the experiment scripts. This left 50 individuals in the final sample for analysis. The mean age of the remaining sample was `r printnum(mean(data$age))` (`r printnum(sd(data$age))`), a majority of participants were female (`r printnum((sum(data$sex == "Female") / nrow(data))*100)`%), and all were white/caucasian without hispanic/Latinx ethnicity. All subjects provided written informed consent in accordance with the Declaration of Helsinki and all procedures were approved by the local Institutional Review Board (Approval #20141014670EP). Each participant received course credit for completing the study.

## Material
### Stimuli
``` {r stimulus_t.test, include = FALSE}
stim.data <- read.xlsx("~/Documents/Nick-Grad/Neta_Lab/depletion_study/study2/Analyses/WorkingMemoryLoads/IAPS_Stim_List.xlsx")
shapiro.test(subset(stim.data, Condition == "POS")$Aro_Mn)
shapiro.test(subset(stim.data, Condition == "NEG")$Aro_Mn)
stim.t.test <- wilcox_test(subset(stim.data, Condition == "POS")$Aro_Mn, subset(stim.data, Condition == "NEG")$Aro_Mn)
```
The stimuli included faces taken from the NimStim [@tottenham_nimstim_2009] and Karolinska Directed Emotional Faces [@lundqvist_karolinska_1998] stimuli sets, as in previous work [@neta_primacy_2010; @brown_cortisol_2017]. The faces consisted of 34 unique identities including 11 angry, 12 happy, and 24 surprised expressions organized pseudorandomly. The scene stimuli were selected from the International Affective Picture System [@lang_international_2008]. A total of 288 scenes (72 positive, 72 negative, and 144 neutral) were selected for the image matrices. The positive and negative images did not differ on arousal (Z = `r printnum(stim.t.test$z_val)`, p = `r printnum(stim.t.test$p.value)`). The scenes were organized into low (two images) and high (six images) cognitive load of either neutral or emotional (equal number of positive and negative) images (Figure 1).

@@

## Procedure
After arriving at the lab, participants provided informed consent prior to completing the task. Participants were randomly assigned to complete one of the task versions, which included 144 [^1] trials split between working memory probe and face rating trials. The trials were self-initiated; the participant clicked a "start" button at the bottom of the screen at the beginning of each trial. After initiating the trial, participants viewed an image matrix of either neutral or emotional images, which the participants were instructed to remember for the duration of the trial. The image matrix was presented for four seconds and the image was either a low or high load matrix consisting of either emotional (equal positive and negative) or neutral images. After the image matrix a happy, angry, or surprised face appeared for one second and the participants rated the face by clicking on either the positive or negative response option. After the face rating, a single image probe appeared, and participants indicated whether or not the image probe was present in the previous image matrix.

@@

# Results
## Subjective ratings
```{r import data and analyze ratings, include = FALSE, echo = TRUE}

### remove psych disorders ### Ask maital?
# data <- data[-c(4, 33, 37, 38, 41, 47),]

@@

### 2019-09-06 Manuscript/Manuscript.tex
Cleaned up scripts
Made a MERP specific data cleaning script...

Starting to polish methods & results.

58 subjects were recruited from the University of Nebraska-Lincoln. All subjects provided written informed consent in accordance with the Declaration of Helsinki and all procedures were approved by the local Institutional Review Board (Approval \#20141014670EP). The data from eight subjects were excluded due to technical difficulties or an error in the experiment script. This left 50 individuals in the final sample.

@@

We used R (Version 3.6.0; {\textbf{???}}) and the R-packages * \}dplyr* {[}@ \}R-dplyr{]}, \emph{broom} (Version 0.5.2; {\textbf{???}}), \emph{forcats} (Version 0.4.0; {\textbf{???}}), \emph{ggplot2} (Version 3.1.1; {\textbf{???}}), \emph{lattice} (Version 0.20.38; {\textbf{???}}), \emph{mousetrap} (Version 3.1.2; {\textbf{???}}), \emph{openxlsx} (Version 4.1.0; {\textbf{???}}), \emph{papaja} (Version 0.1.0.9842; {\textbf{???}}), \emph{plyr} (Version 1.8.4; @ \}R-dplyr; {\textbf{???}}), \emph{purrr} (Version 0.3.2; {\textbf{???}}), \emph{readr} (Version 1.3.1; {\textbf{???}}), \emph{readxl} (Version 1.3.1; {\textbf{???}}), \emph{Rmisc} (Version 1.5; {\textbf{???}}), \emph{stringr} (Version 1.4.0; {\textbf{???}}), \emph{tibble} (Version 2.1.3; {\textbf{???}}), \emph{tidyr} (Version 0.8.3.9000; {\textbf{???}}), and \emph{tidyverse} (Version 1.2.1; {\textbf{???}}) for all our analyses. Data preprocessing was completed in R using the mousetrap package ({\textbf{???}}). First, percent negative ratings were calculated for happy, angry, and surprised faces across all trial types, as well as a percent correct score for the memory probe trials. After, trials were screened for RT outliers. Any trials that were greater than three standard deviations from the mean were removed from the analyses. Additionally, we removed the preceding face rating trial for any incorrect memory probe trials, as these trials can be considered a manipulation failure.

@@

Distributions of ratings were first tested for normality using Shapiro-Wilk's test. The results of all four tests were highly significant (p's \textless{} .001), so non-parametric tests were used for data analysis. Friedman's test results showed significantly different distributions across the conditions \(\chi^{2}\)(3.00) = 27.79, p \textless{} .001. Follow up Wilcoxon signed rank tests revealed that surprise is rated as more negative when holding emotional content in working memory compared to neutral content. Low emotional load ratings were significantly more negative than low, Z = 3.27, p = .001, neutral and high, Z = 3.67, p \textless{} .001, neutral loads. The same was true for high emotional load ratings and low, Z = 4.55, p \textless{} .001, and high, Z = 3.81, p \textless{} .001, neutral loads. However, there was no discernable effect of load. That is, the comparisons between low and high load ratings for both emotional, Z = -1.35, p = .176, and neutral, Z = -0.06, p = .954, load ratings were not significantly different.\footnote{These results are qualitatively the same when analyzing these data with a repeated measures ANOVA.}

@@

\hypertarget{reaction-times}{%
\subsection{Reaction times}\label{reaction-times}}

To assses for normality, the data were first tested with Shapiro-Wilk's test. All conditions appeared to be sampled from normal distributions (p's \textgreater{} .08). A Load (Low, High) X Type (Neutral, Emotional) repeated measures ANOVA was used to assess differences in the RTs. There was a trend towards a main effect of type, F(1, 49) = 2.97, p = 0.09. However, follow up Wilcoxon signed rank tests did not reveal any differences that survived correction for multiple comparisons. There was a trend for RTs of face ratings on high load emotional trials to be longer than high load neutral trials (p =), and a similar pattern for low load emotional trial ratings to take longer than high load neutral trials (p = . One other comparison approached trend levels of signficance, with low load neutral trial ratings taking longer than high neutral loads, (p = . All other comparisons were non-significant (p's \textgreater{} .

\includegraphics{Manuscript_files/figure-latex/plot RT data-1.pdf}

Next, we tested for reaction time differences between positive and negative interpretations of surprised facial expressions within each working memory load condition. Given the non-normality of these data, we used paired-sample Wilcoxon signed rank tests for assessing differences between positive and negative ratings of surprise RTs. This difference was significant \emph{only} for the low emotional load trials, with surprise rated as positive taking signficantly longer than surprise rated as negative (.009). All other comparisons were not significant (p's \textgreater{} .200). Further, this effect survived Bonferroni correction (p = .012).

@@

Finally, we assessed differences in maximum absolute deviation (MAD) across the WM trial conditions. While one of the conditions, low emotional MDs, was not normally distributed (p = .024), all other conditions were normally distributed and repeated-measures ANOVA was used to analyze the MDs across conditions. There was a significant effect of load, F(1.00,196.00) = 5.51, p = .020, such that MDs under high load were larger than trials with low load. There was no significant effect of domain on MDs, F(1.00 196.00) = 0.01, p = .912, nor an interaction of load by domain.
\# Discussion
The effect of high vs.~low load is still not apparent in these data, just like Mattek et al.~2016. An alternative explanation is that the high load manipulation is not sufficiently difficult to recruit the targeted cognitive resources; however, future work will be needed to better test this alternative.

>>>>

Fifty-eight subjects were recruited from the undergraduate research pool at the University of Nebraska-Lincoln. The mean age of the sample was 18.82 (1.19), a majority of participants were female (82.00\%). All subjects provided written informed consent in accordance with the Declaration of Helsinki and all procedures were approved by the local Institutional Review Board (Approval \#20141014670EP). All participants received course credit for completing the study. The data from eight subjects were excluded due to technical difficulties or an error in the experiment script. This left 50 individuals in the final sample.

@@

We used R (Version 3.6.0; {\textbf{???}}) and the R-packages * \}dplyr* {[}@ \}R-dplyr{]}, \emph{BayesFactor} (Version 0.9.12.4.2; {\textbf{???}}), \emph{broom} (Version 0.5.2; {\textbf{???}}), \emph{circlize} (Version 0.4.6; {\textbf{???}}), \emph{coda} (Version 0.19.2; {\textbf{???}}), \emph{cstab} (Version 0.2.2; {\textbf{???}}), \emph{diptest} (Version 0.75.7; {\textbf{???}}), \emph{dotCall64} (Version 1.0.0; {\textbf{???}}; {\textbf{???}}), \emph{fastcluster} (Version 1.1.25; {\textbf{???}}), \emph{fields} (Version 9.8.3; {\textbf{???}}), \emph{forcats} (Version 0.4.0; {\textbf{???}}), \emph{foreach} (Version 1.4.7; {\textbf{???}}), \emph{ggplot2} (Version 3.1.1; {\textbf{???}}), \emph{jpeg} (Version 0.1.8; {\textbf{???}}), \emph{lattice} (Version 0.20.38; {\textbf{???}}), \emph{magrittr} (Version 1.5; {\textbf{???}}), \emph{maps} (Version 3.3.0; {\textbf{???}}), \emph{Matrix} (Version 1.2.17; {\textbf{???}}), \emph{mousetrap} (Version 3.1.2; {\textbf{???}}), \emph{openxlsx} (Version 4.1.0; {\textbf{???}}), \emph{papaja} (Version 0.1.0.9842; {\textbf{???}}), \emph{plyr} (Version 1.8.4; @ \}R-dplyr; {\textbf{???}}), \emph{pracma} (Version 2.2.5; {\textbf{???}}), \emph{processx} (Version 3.3.1; {\textbf{???}}), \emph{psych} (Version 1.8.12; {\textbf{???}}), \emph{purrr} (Version 0.3.2; {\textbf{???}}), \emph{RColorBrewer} (Version 1.1.2; {\textbf{???}}), \emph{Rcpp} (Version 1.0.1; {\textbf{???}}; {\textbf{???}}), \emph{readbulk} (Version 1.1.2; {\textbf{???}}), \emph{readr} (Version 1.3.1; {\textbf{???}}), \emph{readxl} (Version 1.3.1; {\textbf{???}}), \emph{Rmisc} (Version 1.5; {\textbf{???}}), \emph{scales} (Version 1.0.0; {\textbf{???}}), \emph{spam} (Version 2.2.2; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}), \emph{stringr} (Version 1.4.0; {\textbf{???}}), \emph{tibble} (Version 2.1.3; {\textbf{???}}), \emph{tidyr} (Version 0.8.3.9000; {\textbf{???}}), \emph{tidyverse} (Version 1.2.1; {\textbf{???}}), and \emph{yarrr} (Version 0.1.5; {\textbf{???}}) for all our analyses. Data preprocessing was completed in R using the mousetrap package ({\textbf{???}}). First, percent negative ratings were calculated for happy, angry, and surprised faces across all trial types, as well as a percent correct score for the memory probe trials. After, trials were screened for RT outliers. Any trials that were greater than three standard deviations from the mean were removed from the analyses. Additionally, we removed the preceding face rating trial for any incorrect memory probe trials, as these trials can be considered a manipulation failure.

@@

Distributions of ratings were first tested for normality using Shapiro-Wilk's test. The results of all four tests were significant (p's \textless{} .001), so non-parametric tests were used for data analysis. Friedman's test results showed significantly different rank-order distributions across the conditions \(\chi^{2}\)(3.00) = 27.79, p \textless{} .001. Follow up Wilcoxon signed rank tests revealed that surprise is rated as more negative when holding emotional content in working memory compared to neutral content, and this was true for both low and high loads. Low emotional load ratings were significantly more negative than low, Z = 3.27, p = .001, neutral and high, Z = 3.67, p \textless{} .001, neutral loads. The same was true for high emotional load ratings and low, Z = 4.55, p \textless{} .001, and high, Z = 3.81, p \textless{} .001, neutral loads. However, there was no discernable effect of load. That is, the comparisons between low and high load ratings for both emotional, Z = -1.35, p = .176, and neutral, Z = -0.06, p = .954, load ratings were not significantly different.\footnote{These results are qualitatively the same when analyzing these data with a repeated measures ANOVA.}

@@

@@

Finally, we assessed differences in maximum absolute deviation (MAD) across the WM trial conditions. While one of the conditions, low emotional MDs, was not normally distributed (p = .024), all other conditions were normally distributed and repeated-measures ANOVA was used to analyze the MDs across conditions. There was a significant effect of load, F(1.00,196.00) = 5.51, p = .020, such that MDs under high load were larger than trials with low load. There was no significant effect of domain on MDs, F(1.00 196.00) = 0.01, p = .912, nor an interaction of load by domain, F(1.00 196.00) = 0.00, p = .960.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The effect of high vs.~low load is still not apparent in these data, just like Mattek et al.~2016. An alternative explanation is that the high load manipulation is not sufficiently difficult to recruit the targeted cognitive resources; however, future work will be needed to better test this alternative.
